---
title: "Mod 5 Quickfire"
author: "Dr. Cassy Dorff"
date: "2022-09-20"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
```

## Healthcare Outcomes, Perceptions, & Political Ideology

The data set `health` has several variables capturing the health outcomes and politics of U.S. states, such as the infant mortality rate and percent of the population that have a favorable attitude toward the Affordable Care Act. 

## Data dictionary 

To read brief descriptions about the individual variables, see the data dictionary markdown file on github.

## Your task: explore the data

- Be clear. 
- Use R programming only (no python or excel).
- At a minimum aim for two well developed, original different data visualizations (use two different graphical presentations of the data). Avoid making a map.
- Interpret your findings.
- You are *not* allowed to work with other students but you *can* ask me questions.
- Do not cheat and find examples of these data used online. Firstly, that should be rather difficult and will waste your time.  Secondly, be original! 
- You *must* push your work in an .rmd file to github in your *assignments* repo
- You *must* also knit to MD flavored github doc and push this to github too.
- You can edit this document however you want or make your own RMD to use.

## Points

Do not fret! This is not a big percentage of your grade. 

- This is worth 10 points (equivalent of 1 hw) which go into the "participation, quizzes, and assignments" category of your gradebook. 
- 5 points = demonstrated effort but incomplete submission
- 6-8 points = good work, meets expectations, but with some room for improvement.
- 10 points = terrific job, exceeds expectations, little room for improvement.

There are also prizes. 


```{r}
health = read.csv("quickfire_data.csv")
dim(health)
lapply(health, class)
head(health)
```

```{r}
plot1 = ggplot(health, aes(x=percent_supporting_expansion, y=health_score)) +
  labs(title="Health Score by State Based Upon Increased Support for Medicade",
       subtitle = "Comparison between US Republican and Democratic Parties", x = "Percentage Support for Increasing Medicade", y = "Survey Composite Health Scores") +
  geom_jitter(aes(col=gov_party)) +
  scale_color_manual(values = c("Democratic Governor" = "blue", "Repubican Governor" = "red", "Independent Governor" = "green")) +
  geom_smooth(aes(col=gov_party), method="lm", se=F)
plot1
```
The above graph attempts to find a correlation between support for increasing Medicade coverage and the health levels by state and by governor party affiliation. It is apparent by the dispersion of the points across the graph that there is little correlation between the variables. That being said, from the trend line shown for the Republican Governor states, it may be inferred that as the health score of the state decreases, the support for Medicade is increased. This is likely due to the population of states with poor public health infrastructures believing that it may be in their interest to implement a federal health system. 


```{r}
demAvg = health %>% filter(gov_party == "Democratic Governor") %>% summarise(mean(health_score))
demAvg = demAvg[1,1]

repAvg = health %>% filter(gov_party == "Repubican Governor") %>% summarise(mean(health_score))
repAvg = repAvg[1,1]

plot2 = ggplot(health, aes(x=ideology, y=health_score)) +
  labs(title="Health Score compared with per state Ideology and Average Health Score by Party",
       subtitle = "Left = More Liberal, Right = More Conservative. Horizontal Lines = Average Party Health Score", x = "Ideology", y = "Survey Composite Health Scores") +
  geom_point(aes(x=ideology, y=health_score, col=gov_party, size=abs(ideology)), show.legend = FALSE) +
  scale_color_manual(values = c("Democratic Governor" = "blue", "Repubican Governor" = "red", "Independent Governor" = "green")) + geom_hline(yintercept=demAvg, color= "blue") + geom_hline(yintercept=repAvg, color= "red") + geom_vline(xintercept = 0)
plot2 
```
The above graph analyzes how Ideology is correlated with the health score of the state. The graph shows that the more liberal a state is, the higher its health score. This can be seen by the average party health scores and that the blue states which lean conservative almost have all of the lowest health scores among blue states. 
